\chapter{Team 4 Agent Design}\label{team_4_agent_design}

\section{Strategy Overview}
Our overall agent strategy is determined by an evolutionary algorithm which allows our agent to preemptively adapt to an unknown environment. 
The evolutionary algorithm is further complemented by a trust score, which is a measure of our agent's perception of a global community score.
Additionally, the agent is also given a cravings element which simulates real-life humane hunger.

\section{Evolutionary Algorithm}
Our agent strategy began by the developing an evolutionary algorithm which allowed our agent to alternate between various configurations. These 
configurations were determined by an overall trust factor. The agent was trained to determine three configurations: selfish, neutral and selfless.
The selfish nature of the agent was designed for self-prioritisation over community welfare. The selfless configuration was trained to optimise
the global life expectancy whereby, our agent would try to reduce global deaths per iteration. The neutral configuration strikes a balance between
the two previous configurations to try and optimise overall life expectancy inclusive of itself.

\subsection{FoodtoEat}
[Placeholder for FoodtoEat]

\subsection{WaitProbability}
[Placeholder for WaitProbability]

\subsection{Cravings}
[Placeholder for Cravings]

\section{Message Passing}
[Placeholder for Message Passing]

\section{Trust Score}
The trust score was a parameter implemented to try and allow a more dynamic approach built on top of our relatively static evolutionary algorithm.
The trust score meant that our agent was able to build a global trust network based on local interactions. These interactions involve asking other
agents to leave a certain amount of food when we are in the critical state. Another interaction involves our agent evaluating the food that another
agent may be taking.

\section{Treaties}
[Placeholder for Treaties]

\section{Experiments}
[Placeholder for Experiments]